### 2019-07-14 treestats_paper.tex
bigger pop size

\documentclass{article}
\usepackage{fullpage}

@@

>>>>

\documentclass{article}
\usepackage{fullpage}
\usepackage{graphicx}

@@

\paragraph{Example: a selective sweep}
In this view,
site statistics are noisy approximations to the corresponding branch statistic
-- but, how noisy?
How big is the contribution of mutation to the overall sampling variance of a statistic?
Figure~\ref{fig:sweep_duality} shows this in a particular situation:
diversity along the chromosome following a few selective sweeps (two full sweeps and three partial).
The top two plots compare ``branch'' diversity -- i.e., as computed only with tree shape --
to ``site'' diversity computed from sequence generated by 20 independent assignments of mutations to the same tree sequence,
with mutation rates $10^{-9}$ and $10^{-8}$, respectively.
We see that as the mutation rate increases, the signal of decrease in diversity around swept loci becomes more clear,
and approaches ``branch'' diversity.
These were computed using the entire population of 1,000 individuals;
how does sampling variance contribute?
Not much, it turns out -- the bottom plot shows both site and branch diversity
computed from 20 nonoverlapping groups of 100 samples.
Neither site or branch diversity vary much between these samples,
implying that the subsample gives us a good estimate of the whole-population values of each.
However, as we see in the top figure,
whole-population site diversity is itself only a quite noisy estimator of branch diversity.

\begin{figure}
    \includegraphics{{figures/swept.1000.999.1e-09.diversity}.pdf}
    \caption{
        Mean genetic diversity and time to most recent common ancestor
        in 500Kb windows along a 50Mb genome (0.5 Morgans) following several selective sweeps.
        In each case, ``site'' is mean genetic diversity (Tajima's $\pi$) divided by mutation rate,
        and ``branch'' is the corresponding branch statistic described above.
        The tree sequence was produced by simulating mutations under positive selection with mutation rate $10^{-12}$ 
        in a population of size 1000 for 400 generations using SLiM,
        followed by recapitation with $N_e=1000$.
        The selected alleles at the marked sites have selection coefficients between 0.08 and 0.25,
        and are at frequencies 96.8\%, 100\%, 16.25\%, 100\%, and 82.6\% in the final generation, respectively.
        All curves use the same tree sequence, including selected mutations,
        and with additional neutral mutations added.
        \textbf{(top)}
        Diversity within the entire population,
        computed using for mode ``site'' from 20 independent assignments of mutations to the same tree sequence with mutation rate $\mu = 10^{-9}$.
        \textbf{(middle)}
        As in the top panel, but with mutation rate $\mu=10^{-8}$,
        showing that as mutation rate increases, mode ``site'' (divided by $\mu$) converges to mode ``branch''.
        \textbf{(bottom)}
        ``Site'' and ``branch'' diversity within 20 disjoint samples of size 100 each,
        computed on a single tree sequence with mutation rate $\mu = 10^{-9}$.
        \label{fig:sweep_duality}
    }
\end{figure}
